Interim report 2\\ Measles risk assessment, modelling and benefit--cost analysis\\ \vspace{2 mm} {\large David T S Hayman, Tim Carpenter,\\ Jonathan C Marshall, Mick Roberts, Nigel P French}

### 2014-09-01 interimreport2.tex
Figures and cost benefit changes
Large changes to sweave code, especially relating to figures, and
altered c/b ratio analyses after going through with Tim

New Zealand has been working towards elimination of endemic (domestic) measles virus transmission, but has suffered from small, but significant outbreaks of measles after measles introductions from abroad. In this interim report we report the results of statistical analyses of risk factors for measles cases since 2007 in New Zealand during outbreaks, provide updated cost analyses measles outbreaks, and include further modelling of measles outbreaks pre- and post different vaccination scenarios, based on alternative situations. We provide preliminary cost-benefit analyses using the results from those simulations, along with a number of alternative vaccination strategies to achieve different vaccination coverage levels. Our key findings were:
\begin{itemize}
\item text.
\item text.
\item text.
\end{itemize}

@@

As a member of the World Health Organization (WHO) Western Pacific Region, New Zealand is committed to work towards measles elimination, defined as the interruption of endemic (domestic) measles virus transmission, as achieved in the Americas in 2002. A brief review of the history of measles in New Zealand was provided in our previous interim report. In this report were report some regression analyses to determine which populations are most at risk, and the likely outcomes of measles infections based on a number of assumptions, and the cost benefit analyses for vaccination dependent on differing scenarios.

@@

A measles risk assessment has been undertaken by the Ministry of Health to better assess current and future population immunity and high risk groups. Given the current measles outbreak, measles control is a priority for the Ministry and resources are available to control this outbreak and decrease the risk of future outbreaks. In our review of the confidential report to the Western Pacific Regional Verification Commission for Measles Elimination risk assessment provided by the Ministry, titled \emph {Progress Towards Measles Elimination in New Zealand - Final}, found the report to be very thorough, however, we believed additional analyses could further inform the understanding of risk from measles infection. The additional analyses included in this section are multivariate modelling to account for confounding within the univariate analyses.

\subsection{Risk analyses methods}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those data (not shown) suggested that denominator data were required to perform multivariate analyses to avoid confounding results due to a lack of independence among risk factors. Specifically we required $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand to test whether interactions among case covariates provide additional information on risk over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, prioritized ethnicities, and socioeconomic deprivation indices (NZDep) that might exist.

The Otago denominator data provided were not to the same detail as the ESR case data. Notably, the denominator age data was categorised into several classes: 0--5, 6--17, 18--24, 25--64, and 65+ year categories. The denominator ethnicity was not prioritized ethnicity at the Level 1 Ethnic Group Codes, but at the Level 2 Ethnic Group Codes, though with some alternative codes provided. After discussions with Otago we have provided results based on the best available data, though for smaller groups, some results may be unreliable as these are discussed below.

The prioritized ethnicities, therefore, are:  European; Maori; Pacific Peoples; Asian; Middle Eastern/Latin American/African; Other Ethnicity; Residual Categories, though for this report only the first five are used, as these categories covered all the cases. Denominator data were classified according to the recommendations of Otago. Thus, with the 10 NZDep classes and the age classes above, this led us to have 250 categories. Because for measles cases the very young appear to be disproportionately affected (Figure~\ref{fig:agecases}), we split the 0--5 age category into two classes, 0--2 and 3--5 years old, assuming equal numbers of young were born into each age group over the last five years (which is supported by data from NZ statistics (REF)).

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-002}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-004}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

Previous measles analyses, including two in New Zealand by Prof. Roberts, estimated the interruption of measles virus transmission can be achieved by herd immunity when approximately 95 percent of the population is homogeneously immune to measles \citep{roberts0,roberts4}. Thus, while New Zealand immunisation activities have led to measles outbreaks becoming less frequent, with decreasing numbers of cases, outbreaks still occur and the current overall population immunity estimates suggest that approximately 85 to 90 percent of the population is immune to measles, thus the reasons for the ongoing outbreaks are likely due to overall population immunity being less than 95 percent and there being pockets of susceptible, non-immune population remaining. The overall population immunity rate is approximately 89\% and the breakdown of this by age category can be seen in Figure N.

\begin{figure}[h!]
\begin{center}
\begin{Schunk}
\begin{Soutput}
[1] NA
\end{Soutput}
\end{Schunk}
\includegraphics{interimreport2-006}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-008}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

\begin{figure}
     \centering
     \includegraphics[width=0.9\textwidth]{AgeCases.pdf}
     \caption{test}
     \label{fig:agecases}
\end{figure}

This large number of categories lead to both zeroinflation and over-dispersion as there were many categories with zero cases in, particularly in the adult age classes. Furthermore, initial preliminary analyses, including multi and univariate analyses (not shown) suggested little effect of NZDep and several higher order interactions, and therefore we reduced the number of NZDep categories from 1 to 10 to two, 1--5 and 6--10. We also incorporated the 65+ age classes into the 25--64 age category, to make a 25+ age category. By doing so, we reduce the zeroinflation present in the data. Subsequent analyses also suggested that the Middle Eastern/Latin American/African category was over-represented (Figure~\ref{fig:percap}).

\begin{figure}
     \centering
     \includegraphics[width=0.9\textwidth]{PerCapitaCases.pdf}
     \caption{Cases per capita for Prioritized Ethnicities. Note categories are not given, but include Age (0-2, 3-5, 6-17, 18-24, and 25+ in years), NZ deprivation index (1-5, 6-10), and Ethnicity. Points are coloured by Ethnicity.}
     \label{fig:percap}
\end{figure}

However, there are several issues with the data for Middle Eastern, Latin American and African (MELAA) ethnicities, which have small population sizes, and which there were numberous issues with estimating the denominator data for. This category was also an outlier (Figure~\ref{fig:percap}), and thus reducing our ability to make inferences regarding the other classifications and risk factors. Thus, we removed the MELAA category for the below analyses. In the below analyses we used a quasipoission regression model to account for overdispersion in the data. We also account for differences in population sizes by using an offset term, the log(population size). Thus, we used a quasipoisson error structure and the following linear predictor:

\begin{equation} \label{eq:reg}
 log(y) = \alpha + \beta _a (x_a)+ \beta _e(x_e)+ \beta _N (x_N) + \beta _a_e(x_a * x_e)+ log(population)  + \epsilon
  \end{equation}

The numbers of cases per category can be seen in Table N.
\begin{table}\tiny
\caption{Absolute number of measles cases in specific age, ethnicity and socio-economic deprivation categories from 2007-2014}
%latex.default(ttyr, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lllr}
\hline\hline
\multicolumn{1}{c}{NZDep}&\multicolumn{1}{c}{Age}&\multicolumn{1}{c}{Ethnicity}&\multicolumn{1}{c}{Cases}\tabularnewline
\hline
1-5&0-2&Asian&$ 11$\tabularnewline
6-10&0-2&Asian&$  8$\tabularnewline
1-5&3-5&Asian&$  1$\tabularnewline
1-5&6-17&Asian&$ 11$\tabularnewline
6-10&6-17&Asian&$  5$\tabularnewline
1-5&18-24&Asian&$  3$\tabularnewline
6-10&18-24&Asian&$  5$\tabularnewline
1-5&25+&Asian&$ 10$\tabularnewline
6-10&25+&Asian&$ 13$\tabularnewline
1-5&0-2&European&$ 83$\tabularnewline
6-10&0-2&European&$ 64$\tabularnewline
1-5&3-5&European&$ 42$\tabularnewline
6-10&3-5&European&$ 17$\tabularnewline
1-5&6-17&European&$219$\tabularnewline
6-10&6-17&European&$ 80$\tabularnewline
1-5&18-24&European&$ 34$\tabularnewline
6-10&18-24&European&$ 36$\tabularnewline
1-5&25+&European&$ 78$\tabularnewline
6-10&25+&European&$ 51$\tabularnewline
1-5&0-2&Maori&$ 18$\tabularnewline
6-10&0-2&Maori&$ 48$\tabularnewline
1-5&3-5&Maori&$  7$\tabularnewline
6-10&3-5&Maori&$ 11$\tabularnewline
1-5&6-17&Maori&$ 19$\tabularnewline
6-10&6-17&Maori&$ 92$\tabularnewline
1-5&18-24&Maori&$  5$\tabularnewline
6-10&18-24&Maori&$  8$\tabularnewline
1-5&25+&Maori&$  2$\tabularnewline
6-10&25+&Maori&$  6$\tabularnewline
6-10&0-2&MLA&$  3$\tabularnewline
6-10&3-5&MLA&$  1$\tabularnewline
6-10&6-17&MLA&$  1$\tabularnewline
6-10&18-24&MLA&$  2$\tabularnewline
6-10&25+&MLA&$  6$\tabularnewline
1-5&0-2&Pacific&$  5$\tabularnewline
6-10&0-2&Pacific&$ 58$\tabularnewline
6-10&3-5&Pacific&$  3$\tabularnewline
1-5&6-17&Pacific&$  5$\tabularnewline
6-10&6-17&Pacific&$ 22$\tabularnewline
1-5&18-24&Pacific&$  1$\tabularnewline
6-10&18-24&Pacific&$  8$\tabularnewline
1-5&25+&Pacific&$  2$\tabularnewline
6-10&25+&Pacific&$ 11$\tabularnewline
1-5&0-2&None&$  3$\tabularnewline
1-5&3-5&None&$  1$\tabularnewline
1-5&6-17&None&$  3$\tabularnewline
6-10&6-17&None&$  4$\tabularnewline
1-5&18-24&None&$  2$\tabularnewline
6-10&18-24&None&$  1$\tabularnewline
1-5&25+&None&$  5$\tabularnewline
6-10&25+&None&$  3$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:cases}
\end{table}

The numbers of cases per category can be seen in Table N.
\begin{table}\tiny
\caption{Absolute number of measles cases in specific age, ethnicity and socio-economic deprivation categories from 2007-2014}
%latex.default(PerCap, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lllrrr}
\hline\hline
\multicolumn{1}{c}{NZDep}&\multicolumn{1}{c}{Age}&\multicolumn{1}{c}{Ethnicity}&\multicolumn{1}{c}{Popn}&\multicolumn{1}{c}{Cases 1}&\multicolumn{1}{c}{Per Capita}\tabularnewline
\hline
1-5&0-2&Asian&$   6094$&$ 11$&$0.0018$\tabularnewline
6-10&0-2&Asian&$   6806$&$  8$&$0.0012$\tabularnewline
1-5&3-5&Asian&$   6094$&$  1$&$0.0002$\tabularnewline
6-10&3-5&Asian&$   6806$&$  0$&$0.0000$\tabularnewline
1-5&6-17&Asian&$  33918$&$ 11$&$0.0003$\tabularnewline
6-10&6-17&Asian&$  28905$&$  5$&$0.0002$\tabularnewline
1-5&18-24&Asian&$  22917$&$  3$&$0.0001$\tabularnewline
6-10&18-24&Asian&$  34107$&$  5$&$0.0001$\tabularnewline
1-5&25+&Asian&$  96357$&$ 10$&$0.0001$\tabularnewline
6-10&25+&Asian&$  98715$&$ 13$&$0.0001$\tabularnewline
1-5&0-2&European&$  57872$&$ 83$&$0.0014$\tabularnewline
6-10&0-2&European&$  45445$&$ 64$&$0.0014$\tabularnewline
1-5&3-5&European&$  57872$&$ 42$&$0.0007$\tabularnewline
6-10&3-5&European&$  45445$&$ 17$&$0.0004$\tabularnewline
1-5&6-17&European&$ 264330$&$219$&$0.0008$\tabularnewline
6-10&6-17&European&$ 182937$&$ 80$&$0.0004$\tabularnewline
1-5&18-24&European&$ 107649$&$ 34$&$0.0003$\tabularnewline
6-10&18-24&European&$ 117840$&$ 36$&$0.0003$\tabularnewline
1-5&25+&European&$1001916$&$ 78$&$0.0001$\tabularnewline
6-10&25+&European&$ 724317$&$ 51$&$0.0001$\tabularnewline
1-5&0-2&Maori&$  10003$&$ 18$&$0.0018$\tabularnewline
6-10&0-2&Maori&$  30104$&$ 48$&$0.0016$\tabularnewline
1-5&3-5&Maori&$  10003$&$  7$&$0.0007$\tabularnewline
6-10&3-5&Maori&$  30104$&$ 11$&$0.0004$\tabularnewline
1-5&6-17&Maori&$  40461$&$ 19$&$0.0005$\tabularnewline
6-10&6-17&Maori&$ 116640$&$ 92$&$0.0008$\tabularnewline
1-5&18-24&Maori&$  15360$&$  5$&$0.0003$\tabularnewline
6-10&18-24&Maori&$  48495$&$  8$&$0.0002$\tabularnewline
1-5&25+&Maori&$  71217$&$  2$&$0.0000$\tabularnewline
6-10&25+&Maori&$ 192729$&$  6$&$0.0000$\tabularnewline
1-5&0-2&MLA&$    728$&$  0$&$0.0000$\tabularnewline
6-10&0-2&MLA&$   1290$&$  3$&$0.0023$\tabularnewline
1-5&3-5&MLA&$    728$&$  0$&$0.0000$\tabularnewline
6-10&3-5&MLA&$   1290$&$  1$&$0.0008$\tabularnewline
1-5&6-17&MLA&$   2991$&$  0$&$0.0000$\tabularnewline
6-10&6-17&MLA&$   4539$&$  1$&$0.0002$\tabularnewline
1-5&18-24&MLA&$   1710$&$  0$&$0.0000$\tabularnewline
6-10&18-24&MLA&$   3078$&$  2$&$0.0006$\tabularnewline
1-5&25+&MLA&$   8028$&$  0$&$0.0000$\tabularnewline
6-10&25+&MLA&$  10335$&$  6$&$0.0006$\tabularnewline
1-5&0-2&Pacific&$   2093$&$  5$&$0.0024$\tabularnewline
6-10&0-2&Pacific&$  13124$&$ 58$&$0.0044$\tabularnewline
1-5&3-5&Pacific&$   2093$&$  0$&$0.0000$\tabularnewline
6-10&3-5&Pacific&$  13124$&$  3$&$0.0002$\tabularnewline
1-5&6-17&Pacific&$   8541$&$  5$&$0.0006$\tabularnewline
6-10&6-17&Pacific&$  51183$&$ 22$&$0.0004$\tabularnewline
1-5&18-24&Pacific&$   3972$&$  1$&$0.0003$\tabularnewline
6-10&18-24&Pacific&$  22098$&$  8$&$0.0004$\tabularnewline
1-5&25+&Pacific&$  18492$&$  2$&$0.0001$\tabularnewline
6-10&25+&Pacific&$  91533$&$ 11$&$0.0001$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:percap}
\end{table}

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-015}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-016}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-017}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-019}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-020}
\end{center}
\caption{test}
\label{fig:test}
\end{figure}

\section{Regression analyses results}
\label{sub:regression_results}
The results are:

\begin{Schunk}
\begin{Sinput}
> summary(model2)
\end{Sinput}
\begin{Soutput}
Call:
glm(formula = cases ~ Age + Ethnicity + NZDep + Age:Ethnicity + 
    Ethnicity:NZDep + offset(log(Popn)), family = "quasipoisson", 
    data = tp)

Deviance Residuals: 
     Min        1Q    Median        3Q       Max  
-1.98437  -0.58185  -0.00005   0.45759   1.47521

>>>>

New Zealand has been working towards elimination of endemic (domestic) measles virus transmission, but has suffered from small, but significant outbreaks of measles after measles introductions from abroad. In this interim report we present the results of statistical analyses of risk factors for measles cases since 2007 in New Zealand during outbreaks, provide updated cost analyses measles outbreaks, and include further modelling of measles outbreaks pre- and post different vaccination scenarios, based on alternative situations. We provide preliminary benefit--cost analyses using the results from those simulations, along with a number of alternative vaccination strategies to achieve different vaccination coverage levels. Our key findings were:
\begin{itemize}
\item Age is the best predictor of risk of measles infection in multi-variate regression analyses, though some groups such as Pacific people, the less socially deprived, and European and Maori school age children have been more likely to be cases in outbreaks since 2007.
\item Estimates of the proportion of the currently naive New Zealand population requiring additional vaccination to ensure measles does not persist range from 0-53\%, leading to an additional 0--250,000 additional vaccinations.
\item The benefit--cost ratio is dependent on expectations regarding measles transmission: if measles in New Zealand continues to infect the same proportion of secondary cases as in the current 2013--2014 outbreak, additional vaccination is very beneficial, but if outbreaks remained the size of the average outbreaks since 1997, there is no benefit to additional vaccination campaigns.
\end{itemize}

@@

As a member of the World Health Organization (WHO) Western Pacific Region, New Zealand is committed to work towards measles elimination, defined as the interruption of endemic (domestic) measles virus transmission. A brief review of the history of measles in New Zealand was provided in our previous interim report. In this report were present regression analyses to determine which populations are most at risk, the likely outcomes of measles infections within New Zealand based on a number of assumptions, and the benefit--cost analyses for vaccination dependent on differing scenarios.

@@

A measles risk assessment has been undertaken by the Ministry of Health to better assess current and future population immunity and high risk groups. Given the current measles outbreak, measles control is a priority for the Ministry and resources are available to control this outbreak and decrease the risk of future outbreaks. In our review of the confidential report to the Western Pacific Regional Verification Commission for Measles Elimination risk assessment provided by the Ministry, titled \emph {Progress Towards Measles Elimination in New Zealand - Final}, found the report to be very thorough, however, we identified additional analyses that might further inform the understanding of risk from measles infection within New Zealand. The additional analyses included in this section are multivariate modelling to account for confounding within the univariate analyses.

\subsection{Risk analyses methods}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those data (not shown) suggested that denominator data were required to perform multivariate analyses to avoid confounding results due to a lack of independence among risk factors. Specifically we required $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand to test whether interactions among case covariates provide additional information on risk over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, prioritized ethnicities, and socioeconomic deprivation indices (NZDep) that might exist among cases allowing better understanding of risk of measles infection.

The University of Otago denominator data provided were not to the same detail as the ESR case data. Notably, the denominator age data were categorised into several classes: 0--5, 6--17, 18--24, 25--64, and 65+ year categories. The ethnicity denominator data were not Prioritized Ethnicity at the Level 1 Ethnic Group Codes, but at the Level 2 Ethnic Group Codes, though with some alternative codes provided that did not match the Level 2 Ethnic codes. After discussions with the University Otago we have provided results based on the best available data, though for smaller ethnic group categories, some results may be unreliable and these are discussed below.

With the 10 NZDep classes, Prioritized Ethnicities, and the age classes above, this led us to have 250 categories. Because for measles cases the very young appear to be disproportionately affected (Figure~\ref{fig:ageinyears}), we split the 0--5 age category into two classes, 0--2 and 3--5 years old, assuming equal numbers of young were born into each age group over the last five years (which is supported by data from NZ statistics \citep{stats14}).

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-002}
\end{center}
\caption{Age of measles cases in years in New Zealand for two periods, 1997--2014 and 2007--2014}
\label{fig:ageinyears}
\end{figure}

This large number of categories, some with small population sizes, lead to both overdispersion and zeroinflation, as there were many categories with zero cases in, particularly in the adult age classes. Furthermore, initial preliminary analyses, including multi-- and univariate analyses (not shown) suggested little effect of individual NZDep classifications and several higher order interactions, and therefore we reduced the number of NZDep categories from 1 to 10 to two: 1--5 and 6--10. We also incorporated the 65+ age classes into the 25--64 age category, to make a 25+ age category. By doing so, we reduce the zeroinflation present in the data.

The prioritized ethnicities for cases are:  European; Maori; Pacific Peoples; Asian; Middle Eastern/Latin American/African (MELAA); Other Ethnicity; Residual Categories, though for the analyses in this report only the first five are used, as these categories cover the overwhelming number of cases, with only 1.9\% (22/1137) of cases having no Prioritized Ethnicity (Table~\ref{table:case}).

\begin{table}\tiny
\caption{Absolute number of measles cases in specific age, ethnicity and socio-economic deprivation categories from 2007-2014}
%latex.default(ttyr, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lllr}
\hline\hline
\multicolumn{1}{c}{NZDep}&\multicolumn{1}{c}{Age}&\multicolumn{1}{c}{Ethnicity}&\multicolumn{1}{c}{Cases}\tabularnewline
\hline
1-5&0-2&Asian&$ 11$\tabularnewline
6-10&0-2&Asian&$  8$\tabularnewline
1-5&3-5&Asian&$  1$\tabularnewline
1-5&6-17&Asian&$ 11$\tabularnewline
6-10&6-17&Asian&$  5$\tabularnewline
1-5&18-24&Asian&$  3$\tabularnewline
6-10&18-24&Asian&$  5$\tabularnewline
1-5&25+&Asian&$ 10$\tabularnewline
6-10&25+&Asian&$ 13$\tabularnewline
1-5&0-2&European&$ 83$\tabularnewline
6-10&0-2&European&$ 64$\tabularnewline
1-5&3-5&European&$ 42$\tabularnewline
6-10&3-5&European&$ 17$\tabularnewline
1-5&6-17&European&$219$\tabularnewline
6-10&6-17&European&$ 80$\tabularnewline
1-5&18-24&European&$ 34$\tabularnewline
6-10&18-24&European&$ 36$\tabularnewline
1-5&25+&European&$ 78$\tabularnewline
6-10&25+&European&$ 51$\tabularnewline
1-5&0-2&Maori&$ 18$\tabularnewline
6-10&0-2&Maori&$ 48$\tabularnewline
1-5&3-5&Maori&$  7$\tabularnewline
6-10&3-5&Maori&$ 11$\tabularnewline
1-5&6-17&Maori&$ 19$\tabularnewline
6-10&6-17&Maori&$ 92$\tabularnewline
1-5&18-24&Maori&$  5$\tabularnewline
6-10&18-24&Maori&$  8$\tabularnewline
1-5&25+&Maori&$  2$\tabularnewline
6-10&25+&Maori&$  6$\tabularnewline
6-10&0-2&MLA&$  3$\tabularnewline
6-10&3-5&MLA&$  1$\tabularnewline
6-10&6-17&MLA&$  1$\tabularnewline
6-10&18-24&MLA&$  2$\tabularnewline
6-10&25+&MLA&$  6$\tabularnewline
1-5&0-2&Pacific&$  5$\tabularnewline
6-10&0-2&Pacific&$ 58$\tabularnewline
6-10&3-5&Pacific&$  3$\tabularnewline
1-5&6-17&Pacific&$  5$\tabularnewline
6-10&6-17&Pacific&$ 22$\tabularnewline
1-5&18-24&Pacific&$  1$\tabularnewline
6-10&18-24&Pacific&$  8$\tabularnewline
1-5&25+&Pacific&$  2$\tabularnewline
6-10&25+&Pacific&$ 11$\tabularnewline
1-5&0-2&None&$  3$\tabularnewline
1-5&3-5&None&$  1$\tabularnewline
1-5&6-17&None&$  3$\tabularnewline
6-10&6-17&None&$  4$\tabularnewline
1-5&18-24&None&$  2$\tabularnewline
6-10&18-24&None&$  1$\tabularnewline
1-5&25+&None&$  5$\tabularnewline
6-10&25+&None&$  3$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:case}
\end{table}

The numbers of cases per category for the complete data set from 2007-2014 can be seen in Table~\ref{table:percap}. Subsequent regression analyses (not shown) also suggested that the Middle Eastern/Latin American/African category was over-- or underrepresented in per capita rates given the very small sample sizes for this classification (Figure~\ref{fig:percap}, Table~\ref{table:percap}), leading to very large standard error in regression analyses.

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-010}
\end{center}
\caption{Per capita rates of measles infections broken down by ethicity, see Table~\ref{table:percap} for details}
\label{fig:percap}
\end{figure}

\begin{table}\tiny
\caption{Absolute number of measles cases in specific age, ethnicity and socio-economic deprivation categories from 2007-2014}
%latex.default(PerCap, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lllrrr}
\hline\hline
\multicolumn{1}{c}{NZDep}&\multicolumn{1}{c}{Age}&\multicolumn{1}{c}{Ethnicity}&\multicolumn{1}{c}{Population}&\multicolumn{1}{c}{Cases 1}&\multicolumn{1}{c}{Per Capita}\tabularnewline
\hline
1-5&0-2&Asian&$   6094$&$ 11$&$0.0018$\tabularnewline
6-10&0-2&Asian&$   6806$&$  8$&$0.0012$\tabularnewline
1-5&3-5&Asian&$   6094$&$  1$&$0.0002$\tabularnewline
6-10&3-5&Asian&$   6806$&$  0$&$0.0000$\tabularnewline
1-5&6-17&Asian&$  33918$&$ 11$&$0.0003$\tabularnewline
6-10&6-17&Asian&$  28905$&$  5$&$0.0002$\tabularnewline
1-5&18-24&Asian&$  22917$&$  3$&$0.0001$\tabularnewline
6-10&18-24&Asian&$  34107$&$  5$&$0.0001$\tabularnewline
1-5&25+&Asian&$  96357$&$ 10$&$0.0001$\tabularnewline
6-10&25+&Asian&$  98715$&$ 13$&$0.0001$\tabularnewline
1-5&0-2&European&$  57872$&$ 83$&$0.0014$\tabularnewline
6-10&0-2&European&$  45445$&$ 64$&$0.0014$\tabularnewline
1-5&3-5&European&$  57872$&$ 42$&$0.0007$\tabularnewline
6-10&3-5&European&$  45445$&$ 17$&$0.0004$\tabularnewline
1-5&6-17&European&$ 264330$&$219$&$0.0008$\tabularnewline
6-10&6-17&European&$ 182937$&$ 80$&$0.0004$\tabularnewline
1-5&18-24&European&$ 107649$&$ 34$&$0.0003$\tabularnewline
6-10&18-24&European&$ 117840$&$ 36$&$0.0003$\tabularnewline
1-5&25+&European&$1001916$&$ 78$&$0.0001$\tabularnewline
6-10&25+&European&$ 724317$&$ 51$&$0.0001$\tabularnewline
1-5&0-2&Maori&$  10003$&$ 18$&$0.0018$\tabularnewline
6-10&0-2&Maori&$  30104$&$ 48$&$0.0016$\tabularnewline
1-5&3-5&Maori&$  10003$&$  7$&$0.0007$\tabularnewline
6-10&3-5&Maori&$  30104$&$ 11$&$0.0004$\tabularnewline
1-5&6-17&Maori&$  40461$&$ 19$&$0.0005$\tabularnewline
6-10&6-17&Maori&$ 116640$&$ 92$&$0.0008$\tabularnewline
1-5&18-24&Maori&$  15360$&$  5$&$0.0003$\tabularnewline
6-10&18-24&Maori&$  48495$&$  8$&$0.0002$\tabularnewline
1-5&25+&Maori&$  71217$&$  2$&$0.0000$\tabularnewline
6-10&25+&Maori&$ 192729$&$  6$&$0.0000$\tabularnewline
1-5&0-2&MLA&$    728$&$  0$&$0.0000$\tabularnewline
6-10&0-2&MLA&$   1290$&$  3$&$0.0023$\tabularnewline
1-5&3-5&MLA&$    728$&$  0$&$0.0000$\tabularnewline
6-10&3-5&MLA&$   1290$&$  1$&$0.0008$\tabularnewline
1-5&6-17&MLA&$   2991$&$  0$&$0.0000$\tabularnewline
6-10&6-17&MLA&$   4539$&$  1$&$0.0002$\tabularnewline
1-5&18-24&MLA&$   1710$&$  0$&$0.0000$\tabularnewline
6-10&18-24&MLA&$   3078$&$  2$&$0.0006$\tabularnewline
1-5&25+&MLA&$   8028$&$  0$&$0.0000$\tabularnewline
6-10&25+&MLA&$  10335$&$  6$&$0.0006$\tabularnewline
1-5&0-2&Pacific&$   2093$&$  5$&$0.0024$\tabularnewline
6-10&0-2&Pacific&$  13124$&$ 58$&$0.0044$\tabularnewline
1-5&3-5&Pacific&$   2093$&$  0$&$0.0000$\tabularnewline
6-10&3-5&Pacific&$  13124$&$  3$&$0.0002$\tabularnewline
1-5&6-17&Pacific&$   8541$&$  5$&$0.0006$\tabularnewline
6-10&6-17&Pacific&$  51183$&$ 22$&$0.0004$\tabularnewline
1-5&18-24&Pacific&$   3972$&$  1$&$0.0003$\tabularnewline
6-10&18-24&Pacific&$  22098$&$  8$&$0.0004$\tabularnewline
1-5&25+&Pacific&$  18492$&$  2$&$0.0001$\tabularnewline
6-10&25+&Pacific&$  91533$&$ 11$&$0.0001$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:percap}
\end{table}

However, there are several issues with the data for Middle Eastern, Latin American and African (MELAA) ethnicities, and along with small population sizes (Table\ref{table:percap}), and there several issues with estimating the denominator data for this group (University of Otago, personal communication). Thus, we removed this grouping for our subsequent analyses and are left with Asian, European, Maori and Pacific as Prioritized Ethnicities. This left us with 1102/1115 (99\%) of the measles cases with Prioritized Ethnicity recorded from 2007, and 1102/1137 (97\%) of all measles cases recorded since 2007 (Table\ref{table:percap}).

For all our statistical analyses (including those above not shown) we used a Poisson error structure, but in all cases there was a need to account for overdispersion and thus we used, and present, a quasipoission regression model. We also account for differences in population sizes by using an offset term, the log(population size). We used a model simplification approach, by beginning our analyses with all terms and all interactions, and simplifying the models through removal of non-significant higher order interaction terms. Thus, the final model that remained with all significant interaction terms had the following linear predictor:

\begin{equation} \label{eq:reg}
 log(y) = \alpha + \beta _a (x_a)+ \beta _e(x_e)+ \beta _n (x_n) + \beta _a_e(x_a * x_e)+ log(population)  + \epsilon
  \end{equation}
Where $\alpha$ is the intercept, $y$ cases, $_a$ age, $_e$ Prioritized Ethnicity, $_n$ NZDep, and $\epsilon$ the error term.

%\begin{figure}[h!]
%\begin{center}
%<<fig=TRUE,echo=FALSE,width = 11>>=
%hist(tp$cases,xlab="Cases",main='Histogram of cases per category',col='grey',breaks=20)
%@
%\end{center}
%\caption{test}
%\label{fig:test}
%\end{figure}

%\begin{figure}[h!]
%\begin{center}
%<<fig=TRUE,echo=FALSE,width = 11>>=
%plot(tp$cases,main="Cases per category",ylab="Count",pch=16,col="darkgrey")
%@
%\end{center}
%\caption{Cases per capita for Prioritized Ethnicities. Note categories are not given, but include Age (0-2, 3-5, 6-17, 18-24, and 25+ in years), NZ deprivation index (1-5, 6-10), and Ethnicity. Points are coloured by Ethnicity.}
%\label{fig:test}
%\end{figure}
%
%\begin{figure}[h!]
%\begin{center}
%<<fig=TRUE,echo=FALSE,width = 11>>=
%res<-predict(model2)
%plot(exp(res),tp$cases,xlab="results",ylab='predictions',main="Fit",pch=16,col="darkgrey")
%@
%\end{center}
%\caption{test}
%\label{fig:test}
%\end{figure}

%<<echo=FALSE,include=FALSE,results=hide>>=
%cor(exp(res),tp$cases)
%cor.test(exp(res),tp$cases)
%@
%
%\begin{figure}[h!]
%\begin{center}
%<<fig=TRUE,echo=FALSE,width = 11>>=
%hist(model2$residuals,main="Histogram of residuals",xlab="residuals",col="grey")
%@
%\end{center}
%\caption{test}
%\label{fig:test}
%\end{figure}

%\section{Regression analyses results}
%\label{sub:regression_results}
%The results are:

%<<label=regmodel,include=FALSE>>=
%summary(model2)
%@

%<<label=regmodelanova,include=FALSE>>=
%anova(model2,test="F")
%@

\subsection{Regression analyses results}
\label{sub:regression_results}

The distribution of the cases per category used are in Figure~\fig{fig:histcasecat}.
\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-014}
\end{center}
\caption{Histogram of cases per category used in the final regression model}
\label{fig:histcasecat}
\end{figure}

The predicted values plotted against the cases are shown in Figure~\ref{fig:predict}, and the residuals are shown in Figure~\ref{fig:resid}.
\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-015}
\end{center}
\caption{Regression model predictions plotted against the cases (Equation~\ref{eq:reg})}
\label{fig:predict}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics{interimreport2-017}
\end{center}
\caption{Histogram of residuals from the regression model (Equation~\ref{eq:reg})}
\label{fig:resid}
\end{figure}
The significance of the different predictor variables can be seen here, in the ANOVA results:
